\documentclass[conference]{IEEEtran}
\pdfoutput=1

\usepackage[sorting=none,style=ieee,backend=bibtex]{biblatex}
\addbibresource{references.bib}

\usepackage{amsmath}
\usepackage{amsbsy}
\usepackage[noend]{algpseudocode}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{eqnarray}
\usepackage{multicol}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{color}
\usepackage{units}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{vwcol}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{footnote}
\usepackage{blkarray}
\usepackage{comment}

\newtheorem{theorem}{Theorem}
\newtheorem{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{example}{Example}[section]
\captionsetup[table]{position=below}
\def\tablename{Table}
\makesavenoteenv{tabular}
\makesavenoteenv{table}

\allowdisplaybreaks

\ifCLASSINFOpdf

%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\mbf}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\bs}[1]{\ensuremath{\boldsymbol{#1}}}
\newcommand{\mc}[1]{\ensuremath{\mathcal{#1}}}
\newcommand{\cxy}{\ensuremath{C_{X\rightarrow Y}}}
\newcommand{\cyx}{\ensuremath{C_{Y\rightarrow X}}}
\newcommand{\estcxy}{\ensuremath{\hat{C}_{X\rightarrow Y}}}
\newcommand{\estcyx}{\ensuremath{\hat{C}_{Y\rightarrow X}}}
\newcommand{\optcxy}{\ensuremath{C^*_{X\rightarrow Y}}}
\newcommand{\optcyx}{\ensuremath{C^*_{Y\rightarrow X}}}
\newcommand{\fxr}[1]{\ensuremath{f_{X_#1}^{(r)}}}
\newcommand{\fxc}[1]{\ensuremath{f_{X_#1}^{(c)}}}
\newcommand{\fyr}[1]{\ensuremath{f_{Y_#1}^{(r)}}}
\newcommand{\fyc}[1]{\ensuremath{f_{Y_#1}^{(c)}}}
\newcommand{\estfxr}[1]{\ensuremath{\hat{f}_{X_#1}^{(r)}}}
\newcommand{\estfxc}[1]{\ensuremath{\hat{f}_{X_#1}^{(c)}}}
\newcommand{\estfyr}[1]{\ensuremath{\hat{f}_{Y_#1}^{(r)}}}
\newcommand{\estfyc}[1]{\ensuremath{\hat{f}_{Y_#1}^{(c)}}}
\newcommand{\optfxr}[1]{\ensuremath{f_{X_#1}^{(r)*}}}
\newcommand{\optfxc}[1]{\ensuremath{f_{X_#1}^{(c)*}}}
\newcommand{\optfyr}[1]{\ensuremath{f_{Y_#1}^{(r)*}}}
\newcommand{\optfyc}[1]{\ensuremath{f_{Y_#1}^{(c)*}}}
\newcommand{\history}[1]{\ensuremath{\mc{H}_#1}}
\newcommand{\hr}[1]{\ensuremath{\history{i}^{(r)}}}
\newcommand{\hc}[1]{\ensuremath{\history{i}^{(c)}}}
\newcommand{\kl}[2]{\ensuremath{D(#1 \mid \mid #2)}}
\newcommand{\argmin}[1]{\ensuremath{\underset{#1}{\operatorname{argmin}}}}
\newcommand{\argmax}[1]{\ensuremath{\underset{#1}{\operatorname{argmax}}}}
\newcommand{\refclass}{\ensuremath{\tilde{\mc{P}}}}
\newcommand{\refclassr}{\ensuremath{\refclass^{(r)}}}
\newcommand{\refclassc}{\ensuremath{\refclass^{(c)}}}
\newcommand{\mcal}[1]{\ensuremath{\mathcal{#1}}}

%%%%%%%%%%%%%%%%%%%%%%


\begin{document}

\title{A Sample Path Measure of Causal Influence}
%

\author{
Gabriel~Schamberg, \emph{Student Member, IEEE},
Todd~P.~Coleman, \emph{Senior Member,~IEEE}
}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
We present a sample path dependent measure of causal influence between two time series. The proposed measure is a random variable whose expected value is the directed information rate when such a rate exists. A realization of the proposed measure may be used to identify the specific patterns in the data that yield a greater flow of information from one process to another. We demonstrate how sequential prediction theory may be leveraged to obtain accurate estimates of the causal measure at each point in time and introduce a notion of regret for assessing the performance of estimators of the measure. We prove a finite sample bound on this regret that is determined by the regret of the sequential predictors used in obtaining the estimate. We estimate the causal measure for a simulated collection of binary Markov processes using context tree weighting with side information. Finally, given that the measure is a function of time, we demonstrate how estimators of the causal measure may be extended to function in time-varying scenarios.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Granger Causality, KL-Divergence, Sequential Prediction, Markov Chains.
\end{IEEEkeywords}

\IEEEpeerreviewmaketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{01-Introduction}
\input{02-SamplePathMeasure}
\input{03-EstimatingCausalMeasure}
\input{04-DiscreteMarkovChains}
\input{05-Discussion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

\end{document}


